Let’s Disco – Publishing person-level data as Linked Data.

### 2015-01-07 paper/ESWC 2015 Tutorial/ESWC2015Tutorial.tex
ESWC 2015 Tutorial abstract - RDF validation

%
%
\title{Let’s Disco – Publishing person-level data as Linked Data.}
\subtitle{Building a Data Portal with DCAT, DDI-RDF Discovery, PHDD, and other vocabularies.\bz{Ben's Kommentar}
}
%
\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{XXX\inst{1} \and XXX\inst{2}}
%
\authorrunning{XXXXX} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{GESIS – Leibniz Institute for the Social Sciences, Germany\\
\email{\{benjamin,thomas\}}.bosch@gesis.org}

@@

In different real world use cases, the workshop organizers demonstrate the adoption and inclusion of Disco in existing information systems.

@@

@@

Half or full day, including preferences

\tb{we could provide full day if we make the hands on exercises with the SPARQL queries / do you think we should do that?}

@@

Thomas Bosch and Benjamin Zapilko hold a half-day tutorial at the EDDI14 – 6th Annual European DDI User Conference in London\footnote{\url{http://www.eddi-conferences.eu/ocs/index.php/eddi/eddi14/schedConf/program}}.
We got great feedback for this tutorial and had the idea to provide this tutorial at the ESWC conference as well.
At the EDDI conference, we reached the first part of our audience - data professionals and the DDI community.
At the ESWC conference, we want to reach the second part of our audience - Linked Data experts.

>>>>

%
%
\title{Let’s Disco – Publishing person-level data as Linked Data.}
\subtitle{Building a Data Portal with DCAT, DDI-RDF Discovery, PHDD, and other vocabularies.\bz{Ben's Kommentar}
}
%
\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{XXX\inst{1} \and XXX\inst{1} \and Joachim Wackerow\inst{1}}
%
\authorrunning{XXXXX} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{GESIS – Leibniz Institute for the Social Sciences, Germany\\
\email{\{thomas.bosch,benjamin.zapilko,joachim.wackerow\}}@gesis.org}

@@

In different real world use cases, we demonstrate the adoption and inclusion of Disco in existing information systems. 
Existing DDI instances are stored in form of XML documents which can be transformed into DDI-RDF and therefore exposed as Linked Data.
We describe typical constraints on metadata of DDI metadata, as well as how to formulate them by different constraint languages (e.g. OWL 2) and how to actually validate them.

@@

Existing DDI instances are stored in form of XML documents and validated against XML Schemas\footnote{XML Schemas of current DDI version available at \url{http://www.ddialliance.org/Specification/DDI-Lifecycle/3.2/XMLSchema/}}.
DDI-XML documents can be transformed into DDI-RDF and therefore exposed as Linked Data.
For XML, XML Schema is the standard constraint language to validate XML documents.
For RDF, however, there are multiple candidates for languages to express constraints on RDF data - but there is no standard language for RDF validation.
In 2014, two working groups on RDF validation have been established: the W3C RDF Data Shapes working group\footnote{\url{http://www.w3.org/2014/rds/charter}} and the DCMI RDF Application Profiles working group\footnote{\url{http://wiki.dublincore.org/index.php/RDF-Application-Profiles}}.
These working groups address the formulation and validation of RDF constraints.
We will describe typical constraints on metadata of DDI metadata, as well as how to formulate and to validate them.
We will show how to formulate these constraints by different constraint languages such as OWL 2.

@@

We would like to provide a full day workshop as our experience in the previous tutorial about disco showed that a whole day is needed to explain the most important things about disco and to hold the hands-on exercises.
We also want to address the validation of DDI-RDF which we did not covered at the last tutorial.

@@

Thomas Bosch and Benjamin Zapilko hold a half-day tutorial\footnote{tutorial slides avaliable at: \url{http://de.slideshare.net/boschthomas/201412-lets-disco-eddi-2014} and \url{http://de.slideshare.net/boschthomas/201412-lets-disco-2-eddi-2014}} at the EDDI14 – 6th Annual European DDI User Conference in London\footnote{\url{http://www.eddi-conferences.eu/ocs/index.php/eddi/eddi14/schedConf/program}}.
We got great feedback for this tutorial and had the idea to provide this tutorial at the ESWC conference as well.
At the EDDI conference, we reached the first part of our audience - data professionals and the DDI community.
At the ESWC conference, we want to reach the second part of our audience - Linked Data experts.
